Facility and Time Step Discretization Effects in Fuel Cycle Simulation

### 2015-09-18 paper.tex
explain how to measure modeling effects. tweak plots

One difference between the fleet and individual facility modeling is the
\emph{quantized shutdown effect}.  If a fleet-based reactor modeling 3-batch
cores is short a half batch of fuel it will produce $\frac{5}{6}\cdot C$ of
power for that time step where $C$ is the power capacity of a single reactor
of the fleet (e.g. 1000 MWe).  If an individually modeled reactor is short a
half batch of fuel, then it will produce no power that time step.  Time step
duration can also have this effect.  For both fleet and individual reactor
modeling, power goes offline in full time step increments.  For example, if a
3-batch fleet based reactor is short $\frac{1}{2}$ batch of fuel with a 3
month time step, then it will produce $\frac{5}{6} \cdot 3 \text{mo} \cdot C $
of energy.  In contrast, a monthly time step might allow the same reactor to
produce $[\frac{5}{6} \cdot 1 \text{mo} + 2 \text{mo}] \cdot C $ of energy if
the fuel shortage actually ended after only 1 month.

Individually modeled reactors each have their own refueling cycles.  Depending
on how the refueling cycles and outages are staggered, power production can
vary significantly.  For example, if all reactors refuel at the same time,
then all reactors would be offline at the same time.  This is referred to as
the \emph{cycle staggering effect} and does not occur in fleet-based reactor
models.  An example of poor cycle staggering is shown in Figure
\ref{fig:sync-cycle}.  Poor staggering causes spikes of material supply and
demand that can lead to suboptimal resource utilization.  A fuel fabrication
facility might receive requests for new batches for every reactor all at once.
The reactors draw out inventory from the fuel fabrication facility together in
a large quantity of orders.  If the fabrication facility does not have enough
on hand, some number of reactors will need to wait until enough fuel can be
pumped through fabrication.  Avoiding such constraints would require the
fabrication facility to maintain suboptimally large on-hand inventories as a
contingency.

@@

Another difference between fleet-based and individual facility modeling
involves fuel sharing.  Power production by a group of individual reactors for
a fixed amount of fuel shortage may vary depending on how the existing fuel is
distributed.  This is illustrated in Figure \ref{fig:fuel-sharing}.  A system
being short 3 batches might mean that one reactor has no batches and isn't
operating or that 3 reactors are each missing a single batch.  This is
referred to as the \emph{fuel sharing effect}.

@@

One important phenomenon related to time step duration is the \emph{inventory
drawdown effect}.  Increasing time step duration reduces the frequency over
which refueling can occur, resulting in larger impulse drawdown on
inventories. This by itself is not problematic because this is balanced by
correspondingly larger inventory top-up quantities.  The other component of
this effect is the simultaneity of incoming and outgoing material flows on a
given time step.  Within a particular time step facilities do not know about
incoming material when they make commitments for outgoing material.  For
longer time steps, larger incoming quanta of material are not available for
making offers. In general longer time steps and individual reactor modeling
create a need for larger floating inventory buffers in order to avoid
material shortages.

>>>>

One difference between the fleet and individual facility modeling is the
\emph{quantized shutdown effect}.  If a fleet-based reactor modeling 3-batch
cores is short a half batch of fuel it will produce $\frac{5}{6}\cdot C$ of
power for that time step where $C$ is the power capacity of a single reactor
of the fleet (e.g. 1000 MWe).  If an individually modeled reactor is short a
half batch of fuel, then it will produce no power that time step.  Time step
duration can also have this effect.  For both fleet and individual reactor
modeling, power goes offline in full time step increments.  For example, if a
3-batch fleet based reactor is short $\frac{1}{2}$ batch of fuel with a 3
month time step, then it will produce $\frac{5}{6} \cdot 3 \text{mo} \cdot C $
of energy.  In contrast, a monthly time step might allow the same reactor to
produce $[\frac{5}{6} \cdot 1 \text{mo} + 2 \text{mo}] \cdot C $ of energy if
the fuel shortage actually ended after only 1 month.  Quantifying this effect
directly is difficult.  One way of observing it is to compare fleet reactor
scenarios with that have different time step durations (i.e. cases 2 and 4).
Because the fleet reactors have perfect fuel sharing (described below) and no
outage noise, differences observed will likely be due to quantized shutdown.

Individually modeled reactors each have their own refueling cycles.  Depending
on how the refueling cycles and outages are staggered, power production can
vary significantly.  For example, if all reactors refuel at the same time,
then all reactors would be offline at the same time.  This is referred to as
the \emph{cycle staggering effect} and does not occur in fleet-based reactor
models.  An example of poor cycle staggering is shown in Figure
\ref{fig:sync-cycle}.  Poor staggering causes spikes of material supply and
demand that can lead to suboptimal resource utilization.  A fuel fabrication
facility might receive requests for new batches for every reactor all at once.
The reactors draw out inventory from the fuel fabrication facility together in
a large quantity of orders.  If the fabrication facility does not have enough
on hand, some number of reactors will need to wait until enough fuel can be
pumped through fabrication.  Avoiding such constraints would require the
fabrication facility to maintain suboptimally large on-hand inventories as a
contingency.  This effect can be naturally observed as the fluctuations in the
difference between deployed power capacity and generated power.

@@

Another difference between fleet-based and individual facility modeling
involves fuel sharing.  Power production by a group of individual reactors for
a fixed amount of fuel shortage may vary depending on how the existing fuel is
distributed.  This is illustrated in Figure \ref{fig:fuel-sharing}.  A system
being short 3 batches might mean that one reactor has no batches and isn't
operating or that 3 reactors are each missing a single batch.  This is
referred to as the \emph{fuel sharing effect}.  Fleet reactors, by design,
have perfect fuel sharing - all fuel is distributed so that the minimum
possible amount of capacity is offline during shortages. The most correct
measure of fuel sharing inefficiency is the difference between the actual
generated power and the maximum amount of power that could have been generated
among all possible fuel distribution alternatives.  This is difficult to
actually compute. One possible approximation is to count the number of fresh
fuel batches distributed to reactors that ended up not having enough fuel to
operate (i.e. a full core) that time step.  This approximation assumes that
every "wasted" batch of fuel, could have been given to a reactor that only
needed one batch to operate.

@@

One important phenomenon related to time step duration is the \emph{inventory
drawdown effect}.  Increasing time step duration reduces the frequency over
which refueling can occur, resulting in larger impulse drawdown on
inventories. This by itself is not problematic because this is balanced by
correspondingly larger inventory top-up quantities.  The other component of
this effect is the simultaneity of incoming and outgoing material flows on a
given time step.  Within a particular time step facilities do not know about
incoming material when they make commitments for outgoing material.  For
longer time steps, larger incoming quanta of material are not available for
making offers. In general longer time steps and individual reactor modeling
create a need for larger floating inventory buffers in order to avoid material
shortages.  Although direct measures of this effect are difficult, some of its
consequences can be quantified by comparing scenarios with different time step
durations (e.g. cases 2 and 4).
